Blueprint for efficient nuclear spin characterization in color center

### 2024-02-12 article.tex
ready for arxiv

\author{J\"org Wrachtrup}
 \affiliation{3rd Physical Institute University of Stuttgart, Germany}%Lines break 
 \affiliation{Max Planck Institute for solid state physics, Stuttgart, Germany}%Lines break 
 %automatically or can be forced with \\ Third institution, the second for Charlie Author

@@

\begin{abstract}
Nuclear spins in solids offer a promising avenue for developing scalable quantum hardware. Leveraging nearby single-color centers, these spins can be efficiently addressed at the single-site level through spin resonance.
However, characterising individual nuclear spins is quite cumbersome since the utility of characterisation protocols may differ depending on the strength of the hyperfine coupling, necessitating tailored approaches and experimental conditions.
While modified electron spin Hahn echoes like CPMG and XY8 pulse sequences are commonly employed, they encounter significant limitations in scenarios involving spin-1/2 systems, strongly coupled spins, or nuclear spin baths comprising distinct isotopes.
Here, we present a more straightforward approach for determining the hyperfine interactions among each nuclear and the electron spin. 
This method holds promise across diverse platforms, especially for emerging S=1/2 group IV defects in diamond (e.g., SiV, GeV, SnV, PbV) or silicon (T-centre). 
We provide a theoretical framework and adapt it for the use with color-centers exhibiting various spins. Through simulations conducted on nuclear spin clusters, we evaluate different protocols and compare their performance using the Fisher information matrix and Cramer Rao bounds.
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, known as color centers, have been utilized in various quantum applications \cite{awschalom2018quantum}, including quantum networks \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum, meinel2023high}, and quantum registers \cite{bradley2019ten, vorobyov2021quantum, hesselmeier2023measuring, hesselmeier2024high}.

@@

In particular, we consider the following common examples of nuclear spin registers that are ubiquitous in applications. First, the NV-like case with $S=1$ and weakly coupled nuclear spins, operating at high magnetic fields where $\gamma_n B \gg A_{zx}$ which has been extensively studied in the works of \cite{abobeih2019atomic, taminiau2012detection, Zhao:2014aa}.
The capability to observe narrow peaks with analytically predictable positions and contrasts enables the solution of the inverse problem concerning the characterization of the coupling between nuclear spins and electron spins. 

@@

This case is particularly relevant in situations where, in addition to the weakly coupled register, a strongly coupled nuclear spin is utilized, for instance, for repetitive enhancement of the readout.

@@

While the double resonance methods \cite{Bradley:2019aa,zaiser2016enhancing} posses a large promise for nuclear spin spectroscopy, they require additional experimental hardware, and often are not available. 

@@

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Results}
\label{sec:theory}
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/dd_compare2.pdf}
		\caption{Simulated DD signal with 64 $\pi$-pulses for the register reported in  \cite{abobeih2019atomic} consist of 23 nuclear spins, assuming the electron spin is a. spin 1 system b. spin 1/2 system. Simulated Fisher information matrix for the case of the electron spin c. spin 1 system d. spin 1/2 system. The first 23 hyperfine parameters are $A_{zz}$ and the second 23 parameters are $A_{zx}$ of the nuclear spin register.}
		\label{fig:3}
	\end{center}
\end{figure*}
%%%%%%%%%%%%%%%Í%%%%=================
We consider a central spin system of \textit{non-interacting} nuclear spins $I=1/2$ coupled to the central electron spin $\boldsymbol{S}$.
The central electron spin is manipulated resonantly with microwave pulses, which transfer the population between the two sublevels denoted as $m_s=s_0$ and $m_s = s_1$.
These two spin sublevels are separated due to the Zeeman effect and/or zero-field splitting, forming a two-level subsystem with an energy splitting of $\omega_a$.
The Hamiltonian of the system in the secular approximation and in the rotating frame of the applied microwave $\omega_{mw}$ can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is the nuclear Rabi frequency with $\gamma_n^{(k)}$ being the nuclear spin gyromagnetic ratio and $B$ the external magnetic field, and $A_{zz}$ and $A_{zx}$ are the parallel and perpendicular secular components of the hyperfine tensor, respectively.
The Hamiltonian is diagonal in the electron spin subspace, thus the Hamiltonian for the nuclear spins could be rewritten in the electron spin subdomains $s_i$ ($i=0,1$).
The nuclear spin Hamiltonian could be solved for eigenenergies, which determine the precession frequencies of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the eigenvector axes $\vec{n}i =(\frac{s_i A{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$.
It is assumed that the pulse duration $t_p$ is short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible.
For completeness, we start our analysis with the simplest characterization sequence, which is the electron spin Ramsey sequence (free induction decay of the electron).
The Ramsey signal can be obtained as follows:
\begin{gather}
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{j=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(j)}
	\textcolor{gray}{e^{-(\frac{\tau}{T_2^\ast})^m}}
\end{gather}	
The exponential decay is introduced to note the Ramsey signal decays with the electron spin $T_2^\ast$ since the formalism does not account for any relaxation processes. The Ramsey sequence is sensitive to detuning since the electron spin is not refocused; however, it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\

@@

Focusing on weakly coupled nuclear spins or high magnetic field situations, the multiplication over all nuclear spins in Eq. \ref{eq:DD} can be approximated by a summation rule, neglecting the higher-order cross-resonance terms:

@@

Although the Dynamical Decoupling (DD) sequence is well understood and used to characterize nuclear spins, it is not applicable to all systems.
First, DD only works for defects with high magnetic fields and low hyperfine coupling ($\omega_L \gg A_{zz}, A_{zx}$).
Second, the signal is limited to the electron spin $T_2$, which only gradually approaches $T_1$.
Third, signal analysis is rather complicated and time-consuming; one has to collect enough data to ensure that no two nuclear spins overlap in the signal.
Fourth, this sequence does not work for spin 1/2 systems, such as group IV defects in diamond, since the parallel component of the hyperfine coupling cancels their effect on the first order electron spin signal.
Hence, the resonance signal depends on the second order of the perpendicular hyperfine component.
\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/eseem_compare2.pdf}
		\caption{Simulated FFT of ESEEM signal with a zoom into weakly coupled nuclear spins for the register reported \cite{abobeih2019atomic} consist of 23 nuclear spins, assuming the electron spin 1/2 system. b. Enhancing the sensitivity to weakly coupled spins by increasing number of $\pi$-pulses to $N=72$ pulses in each entangling period. $\tau$ is set to the Larmor bright spot. Simulated Fisher information for the case of the electron spin c. one d. one-half system. The first 23 hyperfine parameters are $A_{zz}$ and the second 23 parameters are $A_{zx}$ of the nuclear spin register.}
		\label{fig:ddeseem}
	\end{center}
\end{figure*}

In this work, we consider the 23-nuclear spin cluster characterized in paper \cite{abobeih2019atomic} as a realistic case for our comparison.
First, we apply this to the dynamical decoupling sequence and visualize the obtained results. Fig \ref{fig:3} compares the DD signal for the same nuclear spin register but different defects in diamond.
An intuitive description of the sensitivity of a sequence to the variation of hyperfine coupling could be estimated by considering the width and sensitivity of the position of the resonances.
First, we consider the case of an NV center in diamond. The width of a dip is $\delta \tau = \frac{A_{zx}}{\omega_L^2}$. Taking the derivative of the resonance time, we define the sensitivity of the longitudinal hyperfine coupling as follows, using Eq. \ref{eq:resonance} to simplify:
\begin{gather}
	\delta A_{zz} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_p}{\partial A_{zz}}}} = \frac{2}{\tau_p} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
\end{gather}
If we assume typical values of $T_2^{HE} = 100$ $\mu$s and $\omega_L=500$ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz.
For $S=1/2$, such as group IV defects in diamond, the resonance times are sensitive to the second order of $A_{zx}$ and show no sensitivity to $A_{zz}$.
Hence, even at large $\tau$, the resonances related to nuclear spins will not be well separated.
We define the sensitivity for the spin one-half system as follows:
\begin{gather}
	\delta A_{zx} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
\end{gather}
Assuming typical values of $T_2^{HE} = 100$ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate and approximately 160 times worse than for $S=1$.
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/fig5.png}
		\caption{a. FFT of ESEEM signal for $\tau$ from 10 $\mu$s to 500 $\mu$s with the step of 10 $\mu$s b. 2D correlation of each frequency for different $\tau$.}
		\label{fig:4}
	\end{center}
\end{figure*}

To quantify the difference in sensitivity to estimating the hyperfine parameters, we perform calculations of the Fisher Information Matrix (FIM) for the 23-nuclear spin cluster, with 23 $A_{zz}$ and 23 $A_{zx}$ parameters (see Fig. \ref{fig:3}c,d).
The Fisher Information Matrix is estimated for the probability to measure state $|0\rangle$ and reads as: 
\begin{equation}
	F_{ij}(\textbf{A})=\sum_\tau \frac{\partial p(0, \tau, \textbf{A})}{\partial A_i}\frac{\partial p(0, \tau, \textbf{A})}{ \partial A_j} \frac{1}{p (1-p)}.
\end{equation}
For the typical case of an NV center with $S=1$, since all the nuclear spin resonances are clearly resolved, the Fisher Information Matrix takes a diagonal shape, revealing low covariance between the various nuclear spin resonances.
There is crosstalk between $A_{zz}$ and $A_{zx}$ for a nuclear spin, as the position of the peaks depends on both values.
The Fisher Information provides a bound for the precision of parameter estimation, known as the Cramér–Rao bound: 
\begin{equation}
	\delta A^2 \ge \frac{1}{N} F(A)^{-1}
\end{equation}
The striking difference to the $S=1$ Fisher Information Matrix is the behavior of the defect with $S=1/2$.
First of all, for both spin $S=1/2$ and $S=1$, the sequence's spectral resolution is limited by the $T_1$ relaxation time of the electron spin.
On the other hand, the sequence's sensitivity is limited by $T_2$.
The main principle of 5-pulse ESEEM \cite{kasumaj20085} is to create entanglement (via, e.g., Hahn-echo or CPMG block) before the free evolution of the nuclear spins and then to apply a second correlating sequence afterwards.
In other words, the initial density matrix for the Hahn Echo sequence is modified so that entanglement already exists in the electron polarization terms of the density matrix, thus limited by $T_1$ time.
Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula for this sequence is provided in the appendix \ref{app:5pESEEM}.
During the free evolution time, each nuclear spin oscillates with one of two resonance frequencies obtained from Eq. \ref{eq:H}.

\label{sec:discussion}
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=1.0\textwidth]{pict/cov_compare_vad.pdf}
		\caption{Visualisation of the CramerRao bound for various protocols on existing nuclear spin dataset from \cite{abobeih2019atomic}}
		\label{fig:cov}
	\end{center}
\end{figure*}
Fig. \ref{fig:ddeseem} shows the Fourier transform of the signal.
Zooming into the area close to the Larmor frequency reveals the weakly coupled nuclei.
Each nuclear spin has two peaks in the spectrum referring to $\omega_a$ and $\omega_b$. Expanding the resonance frequencies up to first order with respect to hyperfine coupling shows that nuclear spins appear in order of their coupling strength, making their characterization rather straightforward.
To observe the weakly coupled nuclei more pronouncedly, one can use the DD-ESEEM sequence by adding more $\pi$ pulses in the entangling periods.
To distinguish other nuclei, extra measurements are required.
One method is to use two-dimensional hyperfine correlation spectroscopy (2D Hyscore) \cite{vorobyov2022addressing}.
However, this method is rather time-consuming as two time variables have to be swept.
Optimum time sampling techniques might be used to optimize the timing of 2D sampling.
Here we suggest an alternative way, which is conventionally used in ESR, i.e., sweeping the $\tau$ parameter and keeping track of the frequency amplitude as a function of $\tau$.
Two frequencies that belong to the same nucleus are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$.
Fig. \ref{fig:4} shows two frequencies going to bright and blind spots simultaneously.
Hence, by taking a two-dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nucleus.

@@

Fig. \ref{fig:cov} illustrates the manifestation of the correlation protocol for a spin 1/2 system in terms of the number of nuclear spins one could estimate with each of the protocols.
The dots in the $A_{zz}, A_{zx}$ axes represent the points, with an area bounded by the Cramér-Rao bound around each point.
A nuclear spin is considered detectable (blue points in Fig. \ref{fig:cov}) if the uncertainty of one of the two hyperfine parameters is less than both the absolute value of that parameter and the distance to the closest hyperfine coupling.
Otherwise, it is considered non-identified (shown as red points). 
Additionally, covariance between various parameters is calculated from the off-diagonal elements in the inverse Fisher information matrix.
To illustrate them, an ellipse for each covariance parameter is plotted (colored orange and pink), with two nuclei as the vertices of the ellipse and the small axis size as the Cramér-Rao bound.
This means that in the case of zero covariance between two nuclei, this ellipse turns into a line.
To keep the plot readable, for each nuclear spin, only the ellipse with the largest covariance value is plotted.
These covariances mostly show the crosstalk between different parameters in the presented data and are more strongly present in the DD S=1/2 case. 
The idea behind the ellipse representation is that when the covariance is larger than the self-variance, the spins become indistinguishable.
While this crosstalk is not the limiting factor for informational approach, it might be important when considering a realistic estimator.
It is important to note that this is not strictly the same condition as in experimental identification and is idealized.
We only consider the available information in the plot, assuming that an optimum estimator for the extraction of that information already exists.
In reality, one has to additionally assume a non-ideal estimator procedure, which involves the extraction of the hyperfine parameter values from the raw data and is the subject of further work.
As a result, we see that while for the DD method, a spin 1/2 case for N=16 pulses allows for the identification of only 3 out of 23 nuclear spins, the 5-pulse ESEEM method works with similar success for both S=1 and S=1/2 systems. In total, it is capable of detecting 17-18 out of 23 spins with the constrained measurement time and could be further boosted by using the DD-ESEEM method with multiple pulses in each sensing block analogous to Fig. \ref{fig:ddeseem}.

@@

%\subsection{Advanced filters}
In this work, we conducted a theoretical and numerical comparison of various ESEEM methods for characterizing the nuclear spin clusters around S=1 and S=1/2 type of defects.
We found that for S=1 systems, modified spin echo sequences such as CPMG and XY-N are most suited for single nuclear spin qubit spectroscopy, but for S=1/2 systems, their performance is limited. On the other hand, correlation-type sequences like 3 and 5 pulse ESEEM show more potential for characterizing a diluted nuclear spin bath and perform at least as well as in the S=1 case.
We believe that these methods hold significant potential for preliminary screening of nuclear spin clusters for S=1/2 potential qubit candidate systems, such as G-IV defects in diamond.
Additionally, our method does not require strong microwave pulse fields since the pulses do not need to cope with the nuclear spin Larmor frequency.
While we explored various sensing methods in this work, we did not delve into estimator performance.
We believe that this aspect should be considered in conjunction with adaptive and optimal strategies for controlling the experimental parameters, such as Bayesian Optimal Experimental Design or Machine Learning, to increase the efficiency of estimation.
This could further enhance the characterization capabilities of these methods in practical experimental scenarios.

>>>>

\author{J\"org Wrachtrup}
 \affiliation{3rd Physical Institute University of Stuttgart, Germany}%Lines break 
 \affiliation{Max Planck Institute for solid state research, Stuttgart, Germany}%Lines break 
 %automatically or can be forced with \\ Third institution, the second for Charlie Author

@@

\begin{abstract}
Nuclear spins in solids offer a promising avenue for developing scalable quantum hardware. 
Leveraging nearby single-color centers, these spins can be efficiently addressed at the single-site level through spin resonance.
However, characterising individual nuclear spins is quite cumbersome since the characterisation protocols may differ depending on the strength of the hyperfine coupling, necessitating tailored approaches and experimental conditions.
While modified electron spin Hahn echoes like CPMG and XY8 pulse sequences are commonly employed, they encounter significant limitations in scenarios involving spin-1/2 systems, strongly coupled spins, or nuclear spin baths comprising distinct isotopes.
Here, we present a more straightforward approach for determining the hyperfine interactions among each nuclear and the electron spin. 
This method holds promise across diverse platforms, especially for emerging S=1/2 group IV defects in diamond (e.g., SiV, GeV, SnV, PbV) or silicon (T-centre, P-donors). 
We provide a theoretical framework and adapt it for color-centers exhibiting various spins. Through simulations conducted on nuclear spin clusters, we evaluate different protocols and compare their performance using the Fisher information matrix and Cramer Rao bounds.
\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle
%\tableofcontents
\section{Introduction and background}
Optically active defects in solids, known as color centers, have been utilized in various quantum applications \cite{awschalom2018quantum}, including quantum networks \cite{pompili2021realization}, quantum sensing \cite{degen2017quantum, meinel2023high}, and quantum registers \cite{bradley2019ten, vorobyov2021quantum, hesselmeier2023measuring, hesselmeier2024high, reiner2024high}.

@@

In particular, we consider the following common examples of nuclear spin registers that are ubiquitous in applications. First, the NV-like case with $S=1$ and weakly coupled nuclear spins, operating at high magnetic fields where $\gamma_n B \gg A_{zx}$ which has been extensively studied \cite{abobeih2019atomic, taminiau2012detection, Zhao:2014aa}.
The capability to observe narrow peaks with analytically predictable positions and contrasts enables the solution of the inverse problem concerning the characterization of the coupling between nuclear and electron spins. 

@@

This case is particularly relevant in situations where, in addition to the weakly coupled register, a strongly coupled nuclear spin is utilized, for instance, for repetitive enhancement of readout \cite{neumann2010single}.

@@

While double resonance methods \cite{Bradley:2019aa,zaiser2016enhancing} hold large promise for nuclear spin spectroscopy, they require additional experimental hardware, and often are not available. 

@@

%%%%%%%%%%%%%%%%%%%================= %THEORY
\section{Results}
\label{sec:theory}
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/dd_compare2.pdf}
		\caption{Simulated DD signal with 64 $\pi$-pulses for the register reported in  \cite{abobeih2019atomic} consisting of 23 nuclear spins, assuming the electron spin is a. spin $S=1$ system b. spin $S=1/2$ system. Simulated Fisher information matrix for the case of the electron spin c. spin $S=1$ system d. spin $S=1/2$ system. The first 23 hyperfine parameters are $A_{zz}$ and the second 23 parameters are $A_{zx}$ of the nuclear spin register.}
		\label{fig:3}
	\end{center}
\end{figure*}
%%%%%%%%%%%%%%%Í%%%%=================
We consider a central spin system of \textit{non-interacting} nuclear spins $I=1/2$ coupled to a central electron spin $\boldsymbol{S}$.
The central electron spin is manipulated resonantly with microwave pulses, which transfer the population between the two sublevels denoted as $m_s=s_0$ and $m_s = s_1$.
These two spin sublevels are separated due to the Zeeman effect and/or zero-field splitting, forming a two-level subsystem with an energy splitting of $\omega_a$.
The Hamiltonian of the system in the secular approximation and in the rotating frame of the applied microwave $\omega_{mw}$ can be written as:
\begin{align}
	H = \Delta S_z + \sum_k (\omega_L^{(k)}  + A_{zz}^{(k)} S_z) I_z^{(k)} + A_{zx}^{(k)} S_z I_x^{(k)}
	\label{eq:H}
\end{align}
Where $\Delta = \omega_a - \omega_{mw}$ is the detuning, $\omega_L^{(k)} =\gamma_n^{(k)} B$ is the nuclear Rabi frequency with $\gamma_n^{(k)}$ being the nuclear spin gyromagnetic ratio and $B$ the external magnetic field, and $A_{zz}$ and $A_{zx}$ are the parallel and perpendicular secular components of the hyperfine tensor, respectively.
The Hamiltonian is diagonal in the electron spin subspace, thus the Hamiltonian for the nuclear spins could be rewritten in the electron spin subdomains $s_i$ ($i=0,1$).
The nuclear spin Hamiltonian could be solved for eigenenergies, which determine the precession frequencies of $\omega_i = \sqrt{(\omega_L + s_i A_{zz})^2+ (s_i A_{zx})^2}$, $i=0,1$ along the eigenvector axes $\vec{n}_i =(\frac{s_i A{zx}}{\omega_i}, 0, \frac{\omega_L + s_i A_{zz}} {\omega_i})$.
It is assumed that the pulse duration $t_p$ is short enough ($\omega_L t_p \ll 1$) that the nuclear spin dynamics during this time is negligible.
For completeness, we start our analysis with the simplest characterization sequence, which is the electron spin Ramsey sequence (free induction decay of the electron).
The Ramsey signal can be obtained as follows:
\begin{gather}
	\expval{\sigma_z}_{Ram} = \cos(\Delta \tau) \prod_{j=1}^{n} \Big[ \cos(\frac{\omega_0 \tau}{2}) \cos(\frac{\omega_1 \tau}{2})+ \notag \\
	(\vec{n_0}\cdot \vec{n_1})\sin(\frac{\omega_0 \tau}{2}) \sin(\frac{\omega_1 \tau}{2}) \Big]^{(j)}
	\textcolor{gray}{e^{-(\frac{\tau}{T_2^\ast})^m}}
\end{gather}	
The exponential decay is introduced to note the Ramsey signal decays with the electron spin $T_2^\ast$ since the formalism does not account for any relaxation processes. The Ramsey sequence is sensitive to detuning since the electron spin is not refocused; however, it is possible to sense nuclear spins with vanishing perpendicular hyperfine coupling.\\

@@

Focusing on weakly coupled nuclear spins or high magnetic fields, the multiplication over all nuclear spins in Eq. \ref{eq:DD} can be approximated by a summation rule, neglecting the higher-order cross-resonance terms:

@@

Although the Dynamical Decoupling (DD) sequence is well understood and used to characterize nuclear spins, it is not applicable to all systems.
First, DD only works for defects in high magnetic fields and low hyperfine coupling ($\omega_L \gg A_{zz}, A_{zx}$).
Second, the signal is limited to the electron spin $T_2$, which only gradually approaches $T_1$.
Third, signal analysis is rather complicated and time-consuming; one has to collect enough data to ensure that no two nuclear spins overlap in the signal.
Fourth, this sequence does not work for spin 1/2 systems, such as group IV defects in diamond, since the parallel component of the hyperfine coupling cancels their effect on the first order electron spin signal.
Hence, the resonance signal depends on the second order of the perpendicular hyperfine component.
\begin{figure*}[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/eseem_compare2.pdf}
		\caption{Simulated FFT of ESEEM signal with a zoom into weakly coupled nuclear spins for the register reported \cite{abobeih2019atomic} consist of 23 nuclear spins, assuming the electron spin $S=1/2$ system. b. Enhancing the sensitivity to weakly coupled spins by increasing number of $\pi$-pulses to $N=72$ pulses in each entangling period. $\tau$ is set to the Larmor bright spot. Simulated Fisher information for the case of the electron spin c. $S=1$ d. $S=1/2$ system. The first 23 hyperfine parameters are $A_{zz}$ and the second 23 parameters are $A_{zx}$ of the nuclear spin register.}
		\label{fig:ddeseem}
	\end{center}
\end{figure*}

In this work, we consider the 23-nuclear spin cluster characterized previously \cite{abobeih2019atomic} as a realistic case for our comparison.
First, we apply this to the dynamical decoupling sequence and visualize the obtained results. Fig \ref{fig:3} compares the DD signal for the same nuclear spin register but different defects in diamond.
An intuitive description of the sensitivity of a sequence to the variation of hyperfine coupling could be estimated by considering the width and sensitivity of the position of the resonances.
First, we consider the case of an NV center in diamond. The width of a dip is $\delta \tau = \frac{A_{zx}}{\omega_L^2}$. Taking the derivative of the resonance time, we define the sensitivity of the longitudinal hyperfine coupling, using Eq. \ref{eq:resonance} to simplify:
\begin{gather}
	\delta A_{zz} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_p}{\partial A_{zz}}}} = \frac{2}{\tau_p} \frac{A_{zx}}{\omega_L} \ge \frac{4}{T_2^{HE}} \frac{A_{zx}}{\omega_L}
\end{gather}
If we assume typical values of $T_2^{HE} = 100$ $\mu$s and $\omega_L=500$ kHz, a weakly coupled nuclear spin $A_{zx}=5$ kHz can be distinguished from another weakly coupled nuclear spin with $\delta A_{zz} = 400$ Hz.
For $S=1/2$, such as group IV defects in diamond, the resonance times are sensitive to the second order of $A_{zx}$ and show no sensitivity to $A_{zz}$.
Hence, even at large $\tau$, the resonances related to nuclear spins will not be well separated.
We define the sensitivity for the spin one-half system as follows:
\begin{gather}
	\delta A_{zx} \equiv \frac{\delta \tau}{\abs{\frac{\partial \tau_k}{\partial A_{zx}}}} = \frac{4}{\tau_k} \ge \frac{8}{T_2^{HE}}
\end{gather}
Assuming typical values of $T_2^{HE} = 100$ $\mu$s, a nuclear spin can be distinguished from another one if their transverse hyperfine coupling is separated by $\delta A_{zx} = 80$ kHz, which is quite inaccurate and approximately 160 times worse than for $S=1$.
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=0.9\textwidth]{pict/fig5.png}
		\caption{a. FFT of ESEEM signal for $\tau$ from 10 $\mu$s to 500 $\mu$s with a step of 10 $\mu$s b. 2D correlation of each frequency for different $\tau$.}
		\label{fig:4}
	\end{center}
\end{figure*}

To quantify the difference in sensitivity to estimating the hyperfine parameters, we perform calculations of the Fisher Information Matrix (FIM) for the 23-nuclear spin cluster, with 23 $A_{zz}$ and 23 $A_{zx}$ parameters (see Fig. \ref{fig:3}c,d).
The Fisher Information Matrix is estimated for the probability to measure state $|0\rangle$ and reads as: 
\begin{equation}
	F_{ij}(\textbf{A})=\sum_\tau \frac{\partial p(0, \tau, \textbf{A})}{\partial A_i}\frac{\partial p(0, \tau, \textbf{A})}{ \partial A_j} \frac{1}{p (1-p)}.
\end{equation}
For the typical case of an NV center with $S=1$, since all the nuclear spin resonances are clearly resolved, the Fisher Information Matrix takes a diagonal shape, revealing low covariances between the various nuclear spin resonances.
There is crosstalk between $A_{zz}$ and $A_{zx}$ for a nuclear spin, as the position of the peaks depends on both values.
The Fisher Information provides a bound for the precision of parameter estimation, known as the Cramér–Rao bound: 
\begin{equation}
	\delta A^2 \ge \frac{1}{N} F(A)^{-1}
\end{equation}
The striking difference to the $S=1$ Fisher Information Matrix is the behavior of the defect with $S=1/2$.
First of all, for both spin $S=1/2$ and $S=1$, the sequence's spectral resolution is limited by the $T_1$ relaxation time of the electron spin.
On the other hand, the sequence's sensitivity is limited by $T_2$.
The main principle of 5-pulse ESEEM \cite{kasumaj20085} is to create entanglement (via, e.g., Hahn-echo or CPMG block) before the free evolution of the nuclear spins and then to apply a second correlating sequence afterwards.
In other words, the initial density matrix for the Hahn Echo sequence is modified so that entanglement already exists in the electron polarization terms of the density matrix, thus limited by $T_1$ time.
Fig \ref{fig:seqs} shows the sequence, which can be interpreted as two Hahn Echos separated by a long free evolution $T\gg T_2^\ast$ such that the coherence of the electron spin vanishes. The analytical formula for this sequence is provided in the appendix \ref{app:5pESEEM}.
During the free evolution time, each nuclear spin oscillates with one of two resonance frequencies obtained from Eq. \ref{eq:H}.

\label{sec:discussion}
\begin{figure*}%[htbp]
	\begin{center}
		\includegraphics[width=1.0\textwidth]{pict/cov_compare_vad.pdf}
		\caption{Visualisation of the CramerRao bound for various protocols on an existing nuclear spin dataset from \cite{abobeih2019atomic}}
		\label{fig:cov}
	\end{center}
\end{figure*}
Fig. \ref{fig:ddeseem} shows the Fourier transform of the signal.
Zooming into the area close to the Larmor frequency reveals the weakly coupled nuclei.
Each nuclear spin has two peaks in the spectrum referring to $\omega_a$ and $\omega_b$. Expanding the resonance frequencies up to first order with respect to hyperfine coupling shows that nuclear spins appear in order of their coupling strength, making their characterization rather straightforward.
To observe the weakly coupled nuclei more pronouncedly, one can use the DD-ESEEM sequence by adding more $\pi$ pulses in the entangling periods.
To distinguish other nuclei, extra measurements are required.
One method is to use two-dimensional hyperfine correlation spectroscopy (2D Hyscore) \cite{vorobyov2022addressing}.
However, this method is rather time-consuming as two time variables have to be swept.
Optimum time sampling techniques might be used to optimize the timing of 2D sampling.
Here we suggest an alternative way, which is conventionally used in ESR, i.e., sweeping the $\tau$ parameter and keeping track of the frequency amplitude as a function of $\tau$.
Two frequencies that belong to the same nucleus are correlated because the modulation depth oscillates with blind spot terms of both frequencies $\sin^2(\frac{\omega_0 \tau}{2}) \sin^2(\frac{\omega_1 \tau}{2})$.
Fig. \ref{fig:4} shows two frequencies going to bright and blind spots simultaneously.
Hence, by taking a two-dimensional correlation of the spectrum, one can deduce which two frequencies are correlated and belong to the same nucleus.

@@

Fig. \ref{fig:cov} illustrates the manifestation of the correlation protocol for a spin 1/2 system in terms of the number of nuclear spins one could estimate with each of the protocols.
The dots in the $A_{zz}, A_{zx}$ axes represent the points, with an area bounded by the Cramér-Rao bound around each point.
A nuclear spin is considered detectable (blue points in Fig. \ref{fig:cov}) if the uncertainty of one of the two hyperfine parameters is less than both the absolute value of that parameter and the distance to the closest hyperfine coupling.
Otherwise, it is considered non-identified (shown as red points). 
Additionally, covariance between various parameters is calculated from the off-diagonal elements in the inverse Fisher information matrix.
To illustrate them, an ellipse for each covariance parameter is plotted (colored orange and pink), with two nuclei as the vertices of the ellipse and the small axis size as the Cramér-Rao bound.
This means that in the case of zero covariance between two nuclei, this ellipse turns into a line.
To keep the plot readable, for each nuclear spin, only the ellipse with the largest covariance value is plotted.
These covariances mostly show the crosstalk between different parameters in the presented data and are more strongly present in the DD S=1/2 case. 
The idea behind the ellipse representation is that when the covariance is larger than the self-variance, the spins become indistinguishable.
While this crosstalk is not the limiting factor for informational approach, it might be important when considering a realistic estimator.
It is important to note that this is not strictly the same condition as in experimental identification and is idealized.
We only consider the available information in the plot, assuming that an optimum estimator for the extraction of that information already exists.
In reality, one has to additionally assume a non-ideal estimator procedure, which involves the extraction of the hyperfine parameter values from the raw data. But this is the subject of further work.
As a result, we see that while for the DD method, a spin 1/2 case for N=16 pulses allows for the identification of only 3 out of 23 nuclear spins, the 5-pulse ESEEM method works with similar success for both S=1 and S=1/2 systems. In total, it is capable of detecting 17-18 out of 23 spins within the measurement time constrain. This could be further boosted by using the DD-ESEEM method with multiple pulses in each sensing block analogous to Fig. \ref{fig:ddeseem}.

@@

%\subsection{Advanced filters}
In this work, we conducted a theoretical and numerical comparison of various ESEEM methods for characterizing the nuclear spin clusters around S=1 and S=1/2 type of defects.
We found that for S=1 systems, modified spin echo sequences such as CPMG and XY-N are most suited for single nuclear spin qubit spectroscopy, but for S=1/2 systems, their performance is limited. On the other hand, correlation-type sequences like 3 and 5 pulse ESEEM show more potential for characterizing a diluted nuclear spin bath and perform at least as well as in the S=1 case.
We believe that these methods hold significant potential for preliminary screening of nuclear spin clusters for S=1/2 potential qubit candidate systems, such as G-IV defects in diamond.
Additionally, our method does not require strong microwave pulse fields since the pulses do not need to cope with the nuclear spin Larmor frequency.
While we explore various sensing methods in this work, we did not delve into estimator performance.
We believe that this aspect should be considered in conjunction with adaptive and optimal strategies for controlling the experimental parameters, such as Bayesian Optimal Experimental Design or Machine Learning, to increase the efficiency of estimation.
This could further enhance the characterization capabilities of these methods in practical experimental scenarios.

\section*{Acknowledgments}
We acknowledge financial support by European Union's Horizon 2020 research and innovation program ASTERIQS under grant No. 820394 as well as Spinus, Federal Ministry of Education and Research (BMBF) project MiLiQuant and Quamapolis, Spinning and QR.X the DFG (FOR 2724, INST 41/1109-1 FUGG), the Max Planck Society, and the Volkswagentiftung. M.Z. thanks Max Planck School of Photonics for financial support.
